Big Ideas in Sports Analytics and Statistical Tools for their Investigation

### 2022-07-15 sports_analytics.Rmd
wrote chess paragraph; #6

```{r pkg}
knitr::write_bib(
  c(
    "base",
    "Lahman", 
    "pitchRx", 
    "retrosheet",
    "teamcolors",
    "ggplot2",
    "BradleyTerry2",
    "nflfastR",
    # "nflscrapr", no longer supported, can no longer be installed
    "nflverse",
    "xgboost"
  ),
  file = "pkgs.bib"
)
```

@@

Figure \ref{fig:chess-startup} shows a rendering of the starting chess board obtained via the \pkg{chess} package.

>>>>

```{r pkg}
knitr::write_bib(
  c(
    "base",
    "Lahman", 
    "pitchRx", 
    "retrosheet",
    "teamcolors",
    "ggplot2",
    "BradleyTerry2",
    "nflfastR",
    # "nflscrapr", no longer supported, can no longer be installed
    "nflverse",
    "xgboost",
    "chess", 
    "chessR"
  ),
  file = "pkgs.bib"
)
```

@@

Many tools in sports analytics provide the ability to read, write, and plot data stored in a sport-specific format. 
For example, consider chess, where the sequence of moves in games is often recorded in [Portable Game Notation](https://en.wikipedia.org/wiki/Portable_Game_Notation).
Software tools can then be built around this well-defined format. 
The \pkg{chess} package [@R-chess] provides R users with the ability to read, write, display, and manipulate chess data in PGN format. 
[Application programming interfaces](https://en.wikipedia.org/wiki/API) (APIs) are also common. 
In chess, the \pkg{chessR} package [@R-chessR] allows R users to download game data from the Chess.com API. 
This type of infrastructure, where one package is the "workhorse" that facilitates common generic data operations, and other packages layer on specific functionality, is common in sports analytics.

Figure \ref{fig:chess} shows a rendering of the starting chess board obtained via the \pkg{chess} package, along with the final position in a game recently won by one of the authors (downloaded via the \pkg{chessR} package).
